---
title: "mirai - Shiny Integration"
vignette: >
  %\VignetteIndexEntry{mirai - Shiny Integration}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```
### Shiny Integration

#### Shiny Example Usage

`mirai` may be used as an asynchronous / distributed backend to scale [Shiny](https://shiny.posit.co/) applications and plugs directly into the reactive framework, without the need to use promises.

Depending on the options suppled to `daemons()`, tasks may be distributed across local background processes or multiple networked servers in an efficient and performant manner.

The following example has a button to submit tasks, which will be processed by one of 5 daemons, outputting a pretty spiral pattern upon completion. If more than 5 tasks are submitted at once, the chart updates 5 at a time, limited by the number of available daemons.

```{r shiny, eval=FALSE}
library(mirai)
library(shiny)
library(ggplot2)
library(aRtsy)

# function definitions

run_task <- function() {
  Sys.sleep(5L)
  list(
    colors = aRtsy::colorPalette(name = "random", n = 3),
    angle = runif(n = 1, min = - 2 * pi, max = 2 * pi),
    size = 1,
    p = 1
  )
}

plot_result <- function(result) {
  do.call(what = canvas_phyllotaxis, args = result)
}

status_message <- function(tasks) {
  if (tasks == 0L) {
    "All tasks completed."
  } else {
    sprintf("%d task%s in progress at %s", tasks, if (tasks > 1L) "s" else "", format.POSIXct(Sys.time()))
  }
}

ui <- fluidPage(
  actionButton("task", "Submit a task (5 seconds)"),
  textOutput("status"),
  plotOutput("result")
)

server <- function(input, output, session) {
  # reactive values and outputs
  reactive_result <- reactiveVal(ggplot())
  reactive_status <- reactiveVal("No task submitted yet.")
  output$result <- renderPlot(reactive_result(), height = 600, width = 600)
  output$status <- renderText(reactive_status())
  poll_for_results <- reactiveVal(FALSE)
  
  # create empty mirai queue
  q <- list()

  # button to submit a task
  observeEvent(input$task, {
    q[[length(q) + 1L]] <<- mirai(run_task(), run_task = run_task)
    poll_for_results(TRUE)
  })

  # event loop to collect finished tasks
  observe({
    req(poll_for_results())
    invalidateLater(millis = 250)
    if (length(q)) {
      if (!unresolved(q[[1L]])) {
        reactive_result(plot_result(q[[1L]][["data"]]))
        q[[1L]] <<- NULL
      }
      reactive_status(status_message(length(q)))
    } else {
      poll_for_results(FALSE)
    }
  })
}

app <- shinyApp(ui = ui, server = server)

# mirai setup - 5 local daemons with dispatcher
# switch off cleanup as not necessary (each task is self-contained)
with(daemons(5L, cleanup = FALSE), runApp(app))
```
*Thanks to Daniel Woodie and William Landau for providing the original example on which this is based. Please see <https://wlandau.github.io/crew/articles/shiny.html> which also has examples of the fantastic artwork produced.*

#### Example Using Promises

Alternatively, a 'mirai' may be used seamlessly in any place that accepts 'promises'.

The below example simulates a plot function requiring a long compute in a ‘shiny’ application.

The application starts in around 2s rather than the 8s it would take if not running in parallel.

```{r shinypromises, eval=FALSE}
library(shiny)
library(promises) # for promise pipe

ui <- fluidPage(
  fluidRow(
    plotOutput("one"),
    plotOutput("two"),
  ),
  fluidRow(
    plotOutput("three"),
    plotOutput("four"),
  )
)

make_plot <- function(time) {
  Sys.sleep(time)
  runif(10)
}

args <- list(make_plot = make_plot, time = 2)

server <- function(input, output, session) {
  output$one <- renderPlot(mirai(make_plot(time), .args = args) %...>% plot())
  output$two <- renderPlot(mirai(make_plot(time), .args = args) %...>% plot())
  output$three <- renderPlot(mirai(make_plot(time), .args = args) %...>% plot())
  output$four <- renderPlot(mirai(make_plot(time), .args = args) %...>% plot())
  session$onSessionEnded(stopApp)
}

app <- shinyApp(ui = ui, server = server)

# handle 4 simulateneous computes
with(daemons(4L), runApp(app))

```
*Thanks to Daniel Falbel for providing the original example on which this is based.*
